Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-10-11 Manuscript/manuscript.tex
Update methods

\subsubsection{Simulations of POPC bilayers in aqueous ions}
Simulations of POPC bilayer in pure water or in varying salt concentrations
contained 128 POPC molecules and approximately 50~water molecules per each lipid
in the periodic orthorhombic simulation box.
As a default, water molecules were described by the OPC3 force field~\cite{Izadi16} 
which is currently the most accurate three site rigid water model
\todo{More justification for the choosing the OPC3 water model are needed.
It might be good to show the comparison with the scattering data in bulk water in SI.}. 
In order to test transferability of our newly developed ECC-lipids model, 
we also performed several additional simulations with 
OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005} water models
\todo{The normal TIP3P was tested as well, right?} presented in Supporting Information (SI). 
We used the ECC-ions model for Sodium, Calcium and Chloride ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
Simulations with Lipid14 use ion models by Dang~\cite{smith94,chang1999,dang2006}, and by \AA{}qvist~\cite{aqvist90}. 
MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
The simulation settings used in this work are 
summarized in Table \ref{tbl:mdpar}. 
%and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}.
%\todo{As far as remember, I used there Langevin dynamics instead of thermostated MD, because
%this is done in Amber by default. If this is correct, the information in the
%table do not match with this sentence. Based on semi-extensive testing I made few years
%ago this do not change anything. Anyway, this should be reported consistently.}
Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.

\subsubsection{Simulations of POPC bilayers with cationic surfactants}
An automated topology builder \cite{malde11} was first used to create the structure of
dihexadecyldimethylammonium, which is one of the cationic surfactants used to experimentally quantify the electrometer concept~\cite{scherer89}.
The AmberTools program \cite{amber} was then used to generate the Amber-type force field
parameters. The parameters were converted to the Gromacs format by using
the acpype tool \cite{acpype}. The partial charges were then manually modified
to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
The surfactants were randomly placed among the lipids to form bilayer structures with
mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
6, 14, 21, 35, or 50 surfactants per leaflet. 
%Chloride counter ions were used
%in simulations, because parameters for bromide, which was used in the experiment \cite{scherer89},
%were not available in the standard Gromacs files for Amber force field.
The systems were simulated for 200~ns using Lipid14 model for POPC where reasonable lipid neighbor exchange occurs.
First 20~ns were omitted from the analysis.

The same systems were also simulated with the ECC-lipid model for POPC using the same setup. 
In these simulations the ECC correction was also applied to the cationic surfactant 
by scaling all charges with the same factor as for ECC-lipids, i.e., $f_q=0.8$, 
and by using the atom types with reduced $\sigma$ parameters from ECC-lipids.

@@

%In order to validate the structural properties of the newly developed ECC-lipids model
%we first compared our simulation results without any ions to
%NMR order parameters measurements and x-ray scattering form factors 
%(Fig.~\ref{simVSexpNOions}). 
The x-ray scattering form factor (Fig.~\ref{simVSexpNOions}) and
the area per lipid (Table~\ref{tab:apls}) of POPC bilayer simulated
with the ECC-lipid model are in good agreement with the experimental results,
as well as with the Lipid14 model. Thus, we conclude that the lipid bilayer dimensions
are well captured by the developed ECC-lipid model. 
Areas per lipid of ECC-lipids with various other water models are documented in SI, table~\ref{tab:apls_si}. 
%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.

>>>>

\subsubsection{Simulations of POPC bilayers in aqueous ions}
Simulations of POPC bilayer in pure water or in varying salt concentrations
contained 128 POPC molecules and approximately 50~water molecules per each lipid
in the periodic orthorhombic simulation box.
As a default, water molecules were described by the OPC3 force field~\cite{Izadi16} 
which is currently the most accurate three site rigid water model
\todo{More justification for the choosing the OPC3 water model are needed.
It might be good to show the comparison with the scattering data in bulk water in SI.}. 
In order to test transferability of our newly developed ECC-lipids model, 
we also performed several additional simulations with 
OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005} water models
\todo{The normal TIP3P was tested as well, right?} presented in Supporting Information (SI). 
The ECC-ion models were used for Sodium, Calcium and Chloride ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
Simulation with the Lipid14 model was also ran with the ion model by Dang et al.~\cite{smith94,chang1999,dang2006}.
Simulation data for the Lipid14 model with \AA{}qvist~\cite{aqvist90} ion model
was taken directly from \cite{lipid14POPC0mMNaClfiles}. 
MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
The simulation settings used in this work are 
summarized in Table \ref{tbl:mdpar}. 
%and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}.
%\todo{As far as remember, I used there Langevin dynamics instead of thermostated MD, because
%this is done in Amber by default. If this is correct, the information in the
%table do not match with this sentence. Based on semi-extensive testing I made few years
%ago this do not change anything. Anyway, this should be reported consistently.}
Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.

\subsubsection{Simulations of POPC bilayers with cationic surfactants}
An automated topology builder \cite{malde11} was first used to create the structure of
dihexadecyldimethylammonium.%, which is one of the cationic surfactants used to experimentally quantify the electrometer concept~\cite{scherer89}.
The AmberTools program \cite{amber} was then used to generate the Amber-type force field
parameters. The parameters were converted to the Gromacs format by using
the acpype tool \cite{acpype}. The partial charges were then manually modified
to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
The parameters are available at \cite{??} \todo{To be added}.
The surfactants were randomly placed among the lipids to form bilayer structures with
mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
6, 14, 21, 35, or 50 surfactants per leaflet. Chloride counter ions were used
in simulations, because parameters for bromide, the counterion in the experimental data \cite{scherer89},
were not available in the standard Gromacs files for Amber force field.
The Lipid14 model was used for POPC.
The first 20~ns of the total simulation time of 200~ns
was considered as an equlibration time and was omitted from the analysis.
A reasonable lipid neighbor exchange occured during the simulation.

The same systems were also simulated with the ECC-lipid model for POPC using the same setup.
\todo{Which water model was used in these simulations?}
In these simulations the ECC correction was also applied to the cationic surfactant 
by scaling all charges with the same factor as for ECC-lipids, i.e., $f_q=0.8$, 
and by using the atom types with reduced $\sigma$ parameters from ECC-lipids.

@@

The x-ray scattering form factor (Fig.~\ref{simVSexpNOions}) and
the area per lipid (Table~\ref{tab:apls}) of POPC bilayer simulated
with the ECC-lipid model are in good agreement with the experimental results,
as well as with the Lipid14 model. Thus, we conclude that the lipid bilayer dimensions
are well captured by the developed ECC-lipid model. 
Areas per lipid of ECC-lipids with various other water models are documented in SI, table~\ref{tab:apls_si}. 
%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
